---
title: "Analysis of India DHS Data accounting for sampling weight"
author: "Paul Hsu"
date: "4/1/2025"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning = FALSE)
```
```{r}
install.packages("jtools")
install.packages("svrep")
install.packages("Hmisc")
library(haven)
library("survey")
library("jtools")
library("remotes")
library("svrep")
library(gtsummary)
library(tidyverse)
library(knitr)
library(readr)
library(foreign)
library(plyr)
library(Hmisc)

dhs=read_csv("IAIR7EFL.csv")
#View(dhs)
#generate a subset of the data
dhs.s=subset(dhs,select=c(caseid,v001,v005,v021,v023,v024,v012,v155,v171a,v201,v437,v438,v445))
rm(dhs) #delete dhs data frame from memory
#relabel variables
dhs.s$womenwt=dhs.s$v005/1000000
dhs.s$psu=dhs.s$v021
dhs.s$strata=dhs.s$v023
dhs.s$age=dhs.s$v012
dhs.s$older=(dhs.s$age>30)+0
dhs.s$older=factor(dhs.s$older,levels=c(0,1),labels = c("<=30 yo","> 30 yo"))
dhs.s$literacy=dhs.s$v155
#level needs to match the data
dhs.s$literacy=factor(dhs.s$literacy,
                      levels=seq(0,4),
                      labels=c("Cannot read at all",
                               "Able to read only parts of sentence",
                               "Able to read whole sentence",
                               "No card with required language",
                               "Blind/visually impaired"))
dhs.s$internet.use=(dhs.s$v171a==3)+0
dhs.s$children.born=dhs.s$v201
dhs.s$wt.kg=dhs.s$v437/10
#code missing/unknown as NA
dhs.s$wt.kg[dhs.s$v437>2000]=NA
dhs.s$ht.m=dhs.s$v438/1000
dhs.s$ht.m[dhs.s$v438>2500]=NA
dhs.s$bmi=dhs.s$v445/100
dhs.s$bmi[dhs.s$v445>6000]=NA
```

# Introduction
Before we can start our analyses, we need to use the svydesign function from the “survey” package written by Thomas Lumley. The svydesign function tells R about the design elements in the survey. Once this command has been issued, all you need to do for your analyses is use the object that contains this information in each command. 


# Summary of participant characteristics 
```{r}
#specify the survey design before performing statistical analysis
dhs.d=svydesign(id=dhs.s$caseid, weights=dhs.s$womenwt,strata=dhs.s$strata, data=dhs.s)
#mean age (SD) without accounting for sampling weight
c(mean(dhs.s$age),sd(dhs.s$age))

#mean age (SD) accounting for sampling weight
c(wtd.mean(dhs.s$age,dhs.s$womenwt),wtd.var(dhs.s$age,dhs.s$womenwt)^0.5)

#% of internet use without accounting for sampling weight
mean(dhs.s$internet.use,na.rm=TRUE)

#% of internet use accounting for sampling weight
wtd.mean(dhs.s$internet.use,dhs.s$womenwt,na.rm=TRUE)
```

# Weighted Linear Regression
```{r}
MLR=lm(bmi ~ age+literacy+internet.use+children.born,weights=womenwt, data = dhs.s)
summary(MLR)
```

# Weighted Logistic Regression
```{r}
fit=glm(internet.use~age+literacy+children.born,family=binomial(link="logit"),weights=womenwt,data=dhs.s)
logit.adj=summary(fit)$coefficients
OR.adj=round(cbind(exp(logit.adj[,1]),exp(logit.adj[,1]-1.96*logit.adj[,2]),exp(logit.adj[,1]+1.96*logit.adj[,2]),logit.adj[,4]),3)
colnames(OR.adj)=c("adj OR","Lower 95% CI","Upper 95% CI","p.value")
OR.adj[2,]
```

